---
title: "Match nci-60 DNAm datasets"
author: "K Siegmund"
date: "7/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(limma)
```

# {.tabset}

## Data

Match the dnam data from Spain with the samples analyzed in the USA.

```{r load-data}
load(file = c("data/dnamspain.rda"))
names(dnamspain)
```

```{r ordertissues}
os <- order(dnamspain$targets$tissue,
            dnamspain$targets$title)
dnamspain$targets <- dnamspain$targets[os,]
dnamspain$beta <- dnamspain$beta[,rownames(dnamspain$targets)]
```

What are the characterization data?
```{r tissue}
dnamspain$targets
```

What features from the array are included?
```{r whichfeatures}
table(substr(rownames(dnamspain$beta),1,2))
```
That's all of them.



Let's load the USA data set.
```{r load-usa-data}
load(file = c("data/dnamusa.rda"))
names(dnamusa)
dim(dnamusa$beta)
```

This one doen't have the rs SNPs, and has 61 samples.

```{r ordertissues}
os <- order(dnamusa$targets$tissue,
            dnamusa$targets$title)
dnamusa$targets <- dnamusa$targets[os,]
dnamusa$beta <- dnamusa$beta[,rownames(dnamusa$targets)]
```


```{r vars}
dnamusa$targets
```

Notice that the cell line names are different in the 2 GEO data sets.  I'll need to change these to agree so that I can match the cell lines from the different GEO entries.

## Match samples

```{r match.samples}
midx <- match(dnamusa$targets$title,dnamspain$targets$title)
midx
```

```{r nomatch}
nmidx <- which(is.na(midx))
dnamusa$targets$title[nmidx]
```

Here are the names from Spain, to substitute for the USA names so we can match them.

```{r spainname}
spainname <- c("HS_578T","T47D","COLO-205",
             "HL-60_TB","LOX_IMVI","MDA_N",
             "A549_ATCC","IGR-OV1","NCI-ADR-RES",
             "RC0.1","786-O","RXF393",
             "TK10" ,"U031")
cbind.data.frame(dnamusa$targets$title[nmidx],
                 spainname)             
```

Now substitute these names and find subset common to both sets.
```{r change-name}
dnamusa$targets$title[nmidx] <- spainname
```

Here are the data generated in the USA that are for the same cell lines as previously run in Spain.

```{r match.samples2}
midx <- match(dnamusa$targets$title,dnamspain$targets$title)
midx
```

These don't match:
```{r nomatch2}
nmidx <- which(is.na(midx))
dnamusa$targets$title[nmidx]
```

Now let's subset to common cell lines.
```{r usa}
usa <- dnamusa

usa$targets <- dnamusa$targets[-nmidx,]
usa$beta <- dnamusa$beta[,rownames(usa$targets)]
```

```{r spain}
spain <- dnamspain
midx <- match(spain$targets$title,
              usa$targets$title)
nmidx <- which(is.na(midx))

spain$targets <- spain$targets[-nmidx,]
spain$beta <- spain$beta[,rownames(spain$targets)]
spain$beta <- spain$beta[rownames(usa$beta),]
```

```{r order}
identical(spain$targets$title,usa$targets$title)
```

```{r match-order}
idx<-match(usa$targets$title,spain$targets$title)
spain$targets<- spain$targets[idx,]
spain$beta <- spain$beta[,rownames(spain$targets)]
```

```{r check-order}
identical(spain$targets$title,usa$targets$title)
identical(rownames(spain$targets),colnames(spain$beta))
identical(rownames(usa$targets),colnames(usa$beta))
```

```{r save}
save(usa,file = file.path("data","usa.rda"))
save(spain,file = file.path("data","spain.rda"))
```

```{r sessionInfo}
sessionInfo()
```

