---
title: "Analyze nci-60 pwd"
author: "K Siegmund"
date: "7/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(limma)
library(tidyverse)
```

# {.tabset}

## Data

Load the data sets.
```{r load-data}
load(file = c("data/spain.rda"))
load(file = c("data/usa.rda"))
```

```{r data-list}
names(spain)
dim(spain$beta)
names(usa)
dim(usa$beta)
identical(spain$targets$title,usa$targets$title)
```

Conservation in sample 1.

```{r scattersmooths}
par(mfrow=c(1,2))
smoothScatter(spain$beta[,1],usa$beta[,1],nrpoints=500,
              xlab="Spain",
              ylab="USA",
              main=spain$targets[1,])
a <- (spain$beta[,1] + usa$beta[,1])/2
d <-  spain$beta[,1] - usa$beta[,1]
smoothScatter(a,d,nrpoints=500,
              xlab="Average Beta",
              ylab="Beta-Difference",
              main=spain$targets[1,])
```

```{r conserv}
diffs <- abs(spain$beta - usa$beta)
avgs  <- (spain$beta + usa$beta)/2

pwd <- rowMeans(diffs)
pwa <- rowMeans(avgs)
```

```{r scatsmooths-consv}
smoothScatter(pwa,pwd,nrpoints=500,
              xlab="Average Beta",
              ylab="avg PWD",
              main="59 NCI-60 samples")
```

## Annotate PWD

How many features are in promoter regions?
```{r promoter}
promoter <- ifelse(usa$hm450annot$Regulatory_Feature_Group=="Promoter_Associated",1,0)
sum(promoter)
```

And how many are in enhancer regions?

```{r enhancer}
enhancer <- ifelse(!is.na(usa$hm450annot$Enhancer),1,0)
sum(enhancer)
```

**Stratify by Promoter**

Here is PWD vs Promoter region (yes/no).
```{r boxplots}
boxplot(pwd~promoter)
t.test(pwd~promoter)
```

PWD is lower in Promoters than elsewhere in the genome.

**Stratify by Enhancer**

```{r boxplots-enh}
boxplot(pwd~enhancer)
t.test(pwd~enhancer)
```

We do not find PWD lower in enhancers, like we did Promoters.

**Stratify by 'UCSC_RefGene_Group'**

I'm going to save the first category listed.

```{r ref-gene-gp}
refgenegp <- strsplit(usa$hm450annot$UCSC_RefGene_Group,';')

#  assign feature to the 1st refgene group listed
gp1 <- map(refgenegp,pluck,1)
#  find the null entries (those not gene associated)
gp1.1 <- map_int(gp1,is.null)
#  assign null entries -9 
refgenegp <- unlist(ifelse(gp1.1,-9,gp1))
length(refgenegp)
table(refgenegp)
```

```{r boxplot-refgenegroup}
boxplot(pwd~refgenegp)
```

Let's cut off the top of the figure to spread out the IQR.
```{r boxplot-refgenegroup-lh}
boxplot(pwd~refgenegp,ylim=c(0,.2))
abline(0.05,0,col=2)
```

Next steps:   
1. Bring in other data to characterize the most conserved regions;  
2. Average features local to a region?   
3. Characterize by base context?

## Session Info

```{r sessionInfo}
sessionInfo()
```

